Active learning using the near-Neighborhoods algorithm

### 2024-07-31 old/active-neighborhood-oct.tex
Initial Overleaf Import

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\yh{{\widehat{y}}}
\def\PL{{\rm PL}}
\def\sign{{\rm sign}}

@@

\title{Active learning using the near-neighborhoods algorithm}

@@

\maketitle

\section{Setup}

\begin{itemize}
\item {\it Underlying model of data.}

\begin{itemize}
\item Instances lie in a metric space $(\X, d)$. 
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to an unknown conditional probability function $\Pr(Y|X)$, with
$$ \eta(x) = \E[Y | X=x] = \pr(Y=1|X=x) - \pr(Y=-1|X=x) = 2 \pr(Y=1|X=x) - 1 .$$
\item The {\it Bayes-optimal classifier} on $\X$ is $g^*(x) = \sign(\eta(x))$.
\end{itemize}

\item {\it Task: Finite population labeling.}

\begin{itemize}
\item We are given a finite dataset of $n$ points, and have the
  ability to query the labels of any of them. The response to the
  query $x$ is +1 or -1, chosen according the conditional probability
  function $\eta(x)$. When two or more queries are made on the
  instance $x$, the responses are chosen independently at random
  according to $\eta(x)$, in other words, the labels are not persistant.

\item We are given parameters $0 < \gamma, \delta < 1$.
\item We want a procedure that chooses the next point, or batch of points, to query. This process will be applied repeatedly and forever. At any point the procedure can be asked to make a prediction on any point in $X$. The prediction is $\yh(x) \in \{-1,0,+1\}$. We consider the prediction correct if $\yh(x)=+1$ and $\eta(x)=+1$, if $\yh(x)=-1$ and $\eta(x)=-1$ or $\yh(x)=0$ and $|\eta(x)| < \gamma$. 
\item 
Choosing the queries indepndently at random is guaranteed to reach the correct prediction after querying each example $O(1/\gamma^2)$ times for a total of $O(|X|/\gamma^2)$ queries. Our goal is to reach convergence after much fewer queries.
\item 
Our bounds are pointwise, i.e. we say that the query complexity of the point $x$ is $q(x)$ if for any number of queries larger than $q(x)$ the prediction on $x$ is correct. The hope is that for most $x$, $q(x) \ll |X|/\gamma^2)$.
\item 
Our bounds should be a function of the advantage, as defined in the adaptive KNN paper: $p \gamma^2$
\iffalse
We would like the inferred labels $\yh(x)$ to approach the Bayes-optimal labels $g^*(x)$ as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
\item Specifically, we will then be judged on the number of errors on points $x \in X$ with $|\eta(x)| > \gamma$:
$$ \sum_{x \in X} {\bf 1}(\yh(x) \neq g^*(x),\ |\eta(x)| > \gamma) .$$
The procedure is allowed to fail with probability $\delta$, to account for sampling error.
\fi
\end{itemize}

@@

\begin{figure}[ht!]
\framebox{
\begin{minipage}[t]{6.4in}
\vspace{.1in}
\emph{Variables:}
\begin{itemize}
\item $Q$ (queries made so far)
\end{itemize}

\vspace{.1in}
\emph{Initialization:}
\begin{itemize}
\item Set $Q := \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
\end{itemize}

@@

\vspace{.1in}
\begin{itemize}
\item Query the next unlabeled point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}
\yoav{I am not sure whether the set from which to pick the query
  should exclude $Q$: the queries that have been made o far. if a set
  consists of one point and that point has a weak bias, we need to
  sample the point several times.}
\end{minipage}}
\caption{The sampling procedures. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried. The focused querying procedure also uses threshold $\tau_\ell = 2^{\ell+2}k/n$.} 
\label{alg:sampling}
\end{figure}

\begin{figure}[ht!]
\framebox{
  \begin{minipage}[t]{6.4in}
    \vspace{.1in}
\emph{Variables:}
\begin{itemize}
\item $U_\ell$ The set of points level $\ell$ that are ``known unknowns''
\end{itemize}

\vspace{.2in}
{\bf Active Balls}
\begin{itemize}
\item Initialize uncertainty regions at all levels:
\begin{itemize}
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}

\vspace{.1in}
\item Repeat:
\begin{itemize}
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}
\item Let $\ell'$ be the smallest such level
\item {\tt Focused-query}($\ell', U_{\ell'}$)
\end{itemize}
\item {\tt Background-query}
\item Update $\yh_{\ell}$ values, $\ell \geq 0$, as per equation (\ref{eq:provisional-label})
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = ! \}$
%\yoav{The last run before convergence should contain only disagreements: $!$, not zeros $0$}
\end{itemize}

@@

Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-discrete} in the appendix.

\begin{thm}
Suppose $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c_1}{\sqrt{k}}.$$
\end{enumerate}
\label{thm:large-deviation-bounds}
\end{thm}
\begin{proof}
These follow immediately by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
\end{proof}

@@

\subsection{Mind changes}
\iffalse
\yoav{We say that a prediction {\bf has converged} at level $L_f$ if one of the following holds:
\begin{itemize}
\item For all $\ell\geq L_f$, $\yh_\ell(x) = +1 \mbox{ and } \eta(x) > 0$
\item For all $\ell\geq L_f$, $\yh_\ell(x) = -1 \mbox{ and } \eta(x) < 0$
\item For all $\ell\geq L_f$, $\yh_\ell(x) = 0  \mbox{ and } |\eta(x)| < \gamma$
\end{itemize}

}
\fi

As the level $\ell$ is increased, the label assigned to a data point $x$, which we denote $\yh_{\ell}(x)$, might also change; it may flip between $+1$ and $-1$ a few times, with stretches of $0$ or $!$ in between.

%\yoav{There seems to be a  gap: $\yh_\ell(x)$ is a random variable, but $L_f(x),L_s(x)$ are not.}

We distinguish two critical levels for each $x$: the level $L_s(x)$ beyond which this label no longer flips between $+1$ and $-1$ (although it may be $0$ or $!$), and the level $L_f(x)$ beyond which the label will be correct. See Figure~\ref{fig:two-levels} for a pictorial view.

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{home-stretch.pdf}
\end{center}
\caption{The tentative label $\yh_\ell(x)$ of a given point $x$ can undergo several mind-changes as the neighborhood of $x$ is sampled at successively higher levels. Between levels $L_s(x)$ and $L_f(x)$ there are contradicting labels on $x$, i.e.  $\yh_\ell(x)=!$. Beyond level $L_f(x)$, the correct label will be assigned.}
\label{fig:two-levels}
\end{figure}

We begin with $L_s$.

\begin{defn}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We say $x$ is \emph{obscured} at level $\ell \geq 0$ if the following two conditions hold.
\begin{itemize}
\item No $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) > b_2$.
\item Some $B \in \B_\ell(x)$ has bias $s \cdot \eta_X(B) < -b_1$.
\end{itemize}
Let $L_s(x)$ denote the smallest integer such that $x$ is not obscured at any level $\ell \geq L_s(x)$. If there is no such level, we define $L_s(x) = \infty$.
\label{defn:Ls}
\end{defn}
In words, for levels $\ell < L_s(x)$, there is no ball $B \in \B_\ell(x)$ which has a strong bias towards the correct (Bayes-optimal) label, and there are balls with at least a mild bias towards the wrong label.

Next, we define the level by which the label of a point ought to be apparent.
\begin{defn}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s = \sign(\eta(x))$ be its Bayes-optimal label. We say $x$ is \emph{resolved} at level $\ell \geq 0$ if the following two conditions hold for all $\ell' \geq \ell$.
\begin{itemize}
\item All $B \in \B_{\ell'}(x)$ have $s \cdot \eta_X(B) > -b_1$.
\item Some $B \in \B_{\ell'}(x)$ has $s \cdot \eta_X(B) > b_2$.
\end{itemize}
Let $R_\ell$ denote the points that are resolved by level $\ell$. Let $L_f(x)$ denote the smallest level at which $x$ is resolved, or $\infty$ is there is no such level.
\label{defn:Lf}
\end{defn}
In words, for levels $\ell \geq L_f(x)$, no ball $B \in \B_\ell(x)$ has a strong bias towards the wrong label, and at least one ball has a strong bias towards the correct label.

Thus $L_s(x) \leq L_f(x)$, and there might be a large gap between the two. Before the background sampling reaches level $L_s(x)$, it is quite possible that we will predict the wrong label for $x$. However, once all balls in $\B_{L_s(x)}(x)$ have been adequately sampled, point $x$ will either (i) be assigned label $0$ or $!$, and put in the uncertainty set, in which case it will benefit from focused sampling, or (ii) will get the correct label. Once the balls in $\B_{L_f(x)}(x)$ have been sampled, condition (ii) will hold.

We define the set $F_\ell$ or include all of the points $x \in X$ such that $L_s(x) \leq \ell \leq L_f(x)$. The probability of $F_\ell$ determines the sample complexity of the active learning algorithm

\subsection{A possible characterization of $L_f(x),L_s(x)$}

\renewcommand{\P}{{\cal P}}
We are given a distribution over $\P$ over $\X \times \Y$ and a point $x \in X$.
We sample a set of unlabeled examples $X \subset \X$, and then run our active learning algorithm.

Our goal is to upper bound the number of samples $|\X|$ and the number
of queries the algorithm needs before the prediction of the algorithm
on $x$ converges to the correct label (and will not change with additional samples).

As we are dealing with the true distribution, we redifine $\B_\ell(x)$
to be all balls containing $x$ whose probability is in the range
$\left[ \frac{1}{2^{\ell-1}}, \frac{1}{2^\ell}\right]$.

Given the distribution $\P$, an edge parameter $\gamma>0$, a
prediction point $x$ and a level $\ell$ we partition the balls in $\B_{\ell}(x)$ into three 
subsets:
$$\B_{\ell}^+ = \{ B \in \B_\ell(x) | \eta(B)  > \gamma\},\;\;
\B_{\ell}^- = \{ B \in \B_\ell(x) | \eta(B)  < -\gamma\},\;\;
\B_{\ell}^0 = \{ B \in \B_\ell(x) | -\gamma \leq \eta(B) \leq \gamma\}$$

let the ``correct label'' of x be $s$, i.e. $s = \sign(\eta(x))$.
We define the ``true'' values of  $L_f(x),L_s(x)$ as follows:
\begin{itemize}
\item $L_s(x)$, is the smallest integer $m$ such that for all $\ell\geq
  m$, $\B_\ell^s \neq \emptyset$ while $\B_\ell^{-s} = \emptyset$.
\item $L_f(x)$ is the smallest integer $n \leq L_s(x)$ such that for
  all $n\leq \ell < L_s(x)$, $\B_\ell^s \neq \emptyset$ and $\B_\ell^{-s} \neq \emptyset$.
\end{itemize}

In other words the level $L_s(x)$ is the minimal level below which the
prediction is always correct. $L_f(x)$ is the minimal level above $L_s(x)$
such that the prediction on $x$ for the levels between  $L_f(x)$ and
$L_s(x)$ is always $!$.

\subsection{Another definition of two critical levels}
\yoav{I prefer to work with this definition}

For each point $x \in X$, we define two critical levels $L_0(x)$ and $L_1(x)$.

Let $s = \mbox{sign}(\eta(x)) \in \{-1, +1\}$. Recall that $\PL_\ell(x) \subseteq \{-1,+1\}$ denotes the ``possible labels'' for $x$ after balls in $\B_\ell(x)$ have been queried.

We will define $L_0(x)$ and $L_1(x)$ so that: $s$ is guaranteed to be in $\PL_\ell(x)$ for all $\ell \geq L_0(x)$ and $-s$ is guaranteed to \emph{not} be in $\PL_\ell(x)$ for all $\ell \geq L_1(x)$. More precisely:
\begin{itemize}
\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$,
\begin{itemize}
\item there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$, and
\item any ball $B' \in \B_{\ell'}(x)$ with $B' \subset B$ has $s \cdot \eta_X(B') > -b_1$.
\end{itemize}
\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, any ball $B \in \B_{\ell'}(x)$ either has $s \cdot \eta_X(B) > -b_1$ or contains some $B' \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B') > b_2$.
\item \yoav{I would leave the definition of $L_0$ unchanged, but define $L_1$ as follows} $L_1(x)$ is the smallest level $\ell$ such that for all $L_0(x) > \ell' \geq \ell$, 
there exists two balls $B^+,B^- \in \B_{\ell'}(x)$ such that $\eta_X(B^+) > b_2$ and $\eta_X(B^-) < -b_2$.
\end{itemize}

Without containment:
\begin{itemize}
\item $L_0(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is a ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) > b_2$.
\item $L_1(x)$ is the smallest level $\ell$ such that for all $\ell' \geq \ell$, there is no ball $B \in \B_{\ell'}(x)$ with $s \cdot \eta_X(B) < -b_1$.
\end{itemize}

\section{Characterizing $S_L,S_F$  for Euclidean balls}
Suppose the input
distribution is uniform on a ball $B$ of radius one  around the
origin. Denote the ``positive'' set to be $C$. Let $\gamma>0$ be the
edge such that
\[
  P(y=1|X=x) =\begin{cases}
    1/2+\gamma & \mbox{ if } x \in C \\
    1/2-\gamma & \mbox{ if } x \notin C
    \end{cases}
  \]

  $L_f(x)$ corresponds to half the distance between $x$ and the boundary

  $L_s(x)$ can be bounded independently of $x$ if $r$ is the largest
  radius such that both $C$ and $C^c$ can be expressed as a union of
  balls of radius $r$.

  We can define a bondary of bounded convexity in this way.

 For other definitions of ``ball'' we can use a similar
 characterization: $L_s(x)$ corresponds to the smallest ball
 containing both $x$ and a point with the opposite label. $L_f(x)$
 corresponds to the largest radius such that both $C$ and $C^c$ can be
 written as a union of balls.

\section{Examples}
\begin{itemize}
    \item 1D
    \begin{itemize}
        \item Single threshold with jump in $\eta$ ($-c$ on one side, $+c$ on the other side)
        \item Single threshold with more general $\eta$, e.g. Tsybakov condition
        \item Multiple thresholds with $-c, +c$
    \end{itemize}
    \item 2D all Euclidean balls
    \begin{itemize}
        \item Linear separator with $-c, +c$, e.g. uniform distribution over $x$\\
        I this case $L_s(x)$ is lower bounded the the probability of a
        ball whose radius is half the distance between $x$ and the
        linear separator. $L_f(x)$ is close to 0. Two large balls, one
        shifted towards the positive, the other shifted towards the
        negative, will suffice to create a disagreement.

        \item Separator of bounded curvature on both sides
    \end{itemize}
    \item Dyadic tree: instead of limiting attention to boxes that are cells of the tree, we look at all boxes of all sizes that supported on the underlying gridding of space. We get better bounds for two reasons:
    \begin{itemize}
        \item We look at large boxes, not just leaves of the tree.
        \item We use all boxes, not just those that are nested like the tree; this makes the boundary apparent earlier using !.
    \end{itemize}
    \item Arbitrary dimension, but Holder-smoothness and Tsybakov margin conditions (usual setting in nonparametric estimation). The rates obtainable using passive random sampling were determined by Audibert-Tsybakov. For active, can look at Hanneke. Will find out.
    \item A one-dimensional example, e.g. with data points in [0,1] and with
$\eta > 3/4$ or $\eta < 1/4$ everywhere. In this case, the balls could
be all intervals, or perhaps a finite collection of them based on some
gridding of the line. The Bayes-optimal classifier would have a finite
number of sign-changes, and we could quantify which of these we'd pick
up by each level of sampling.

\item A high-dimensional Euclidean setting in which we use balls
corresponding to the cells of a tree (e.g. RP tree).

\item A high-dimensional metric space setting in which we have a
collection of metric balls and we give label complexity bounds based
on smoothness conditions (Holder-smoothness of the $\eta$ function)
and boundary conditions (like Tsybakov margin condition).

@@

\caption{\yoav{Picture should be changed to include circles containing $x$ but not centered at $x$.} This figure provides some examples for the concept of mind changes. Here we consider true probabilities, rather than estimates. The bounding rectangle defines the input space $X$. The blue region indicate $x$'s where the probability of +1 is larger than that of $-1$, the red region indicates the revers polarity. More specifically, we assume that in the blue region the probability of +1 is 0.6, while in the red region the probability of +1 is 0.4.
The black dots correspond to points on which we want to make a prediction. We consider all circles centered at each point. The {\em bias} of  circle is the probability of +1 conditioned on the point being in the circle. As labels are accumulated, the bias of smaller and smaller circles can be determined. For any point there is a large enough circle that contains the whole input space. We assume that the overall bias is negative, hence for a sufficiently large circle the bias is negative. This is indicated by the dotted circle which is always the largest circle. The solid line circles are the largest circles with the polarity indicated by the color of the line. The smallest circle corresponds to the last mind change which we also call the convergence circle.
A point is considered "easy" if its convergence circle is large, because that means that a small number of queries is required to find the correct label.
Following that definition we get that the order of points from the easiest to the hardest is {\bf a,b,c,e,d}. Comparing e and d we see that d has a smaller convergence circle, and one mind change, which (e) has a larger convergence circle, but has 4 mind changes.
}

@@

>>>>

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\yh{{\widehat{y}}}
\def\PL{{\rm PL}}

@@

\title{Active learning using the near-Neighborhoods algorithm}

@@

\maketitle
\section{Setup}

\begin{enumerate}
\item {\it Pool-based active learning.}

\begin{itemize}
\item We have a finite dataset $X$ of $n$ points, and we are able to query the labels of any of them.
\item We want a procedure that chooses the next point, or batch of points, to query. This process might be stopped at any time, and then the remaining labels will need to be inferred.
\item The goal is for the inferred labels to converge to their ``gold-standard'' values (to be defined) as quickly as possible, ideally much faster than would occur if the query points were chosen uniformly at random.
\end{itemize}

\item {\it Model of data.}

\begin{itemize}
\item The dataset $X$ consists of $n$ independent samples from a fixed but unknown distribution $\mu$ on a metric space $(\X, d)$. 
\item Label space $\Y = \{-1,+1\}$. Each $x \in \X$ is labeled according to conditional probability function $\eta(x) = \pr(Y=1|X=x)$, also unknown.
\item The Bayes-optimal classifier on $\X$ is $g^*(x) = \mbox{sign}(2 \eta(x) - 1)$. This is the gold-standard for inferred labels.
\end{itemize}

s\section{Deterministic algorithm}

\begin{enumerate}
\item Define the highest level to be $l_{max} = \log_2 n$.
\item We have a collection of balls $\B$, organized into levels $\B_\ell$, where $\ell = 0, 1, 2, \ldots$. $\B_\ell$ contains all balls whose probability (wrt the uniform distribution over the examples) satisfies $\frac{1}{2^l} \leq p(B) < \frac{1}{2^{l-1}}$
\item We define $\eta(B) = \pr(Y=1|X \in B)$,
%We will use the shorthand
%$$ \B_{\leq \ell} = \bigcup_{\ell' \leq \ell} \B_{\ell'} $$
%to denote balls upto and including level $\ell$.

\item Let $\B(x) \subset \B$ denote the balls that may be consulted in determining the label of point $x$. These could, for instance, be all balls that contain $x$. Or they could be all balls centered at some $x \in X$

\item The algorithm can query an oracle for a (of labels) for a set of balls $Q_\ell \subset \B_\ell$. The oracle responds with a set of data points such that there are at least $k$ independent samples in each ball $B \in Q_\ell$.

\item
Using these samples we can label each ball in $Q_\ell$ as $y(B)=+1/0/-1$ such that, with probability of $Q_\ell$ such that with probability $1-\delta$ and for some $\epsilon>0$ the following holds for each $B \in Q_\ell$: 
\begin{itemize}    
\item $y(B) = +1 \implies \eta(B) \geq 1-\epsilon$
\item $y(B) = -1 \implies \eta(B) \leq \epsilon$
\item $y(B) = 0 \implies |\eta(B)-1/2| \leq \epsilon$
\end{itemize}

\item Once we have queried all balls in $\B_{\ell}(x)$, we determine the set of possible labels of $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$. It is defined as follows:
\begin{itemize}
\item $\PL_\ell(x)$ contains $+1$ iff there exists $B \in \B_{\ell}(x)$ with $y(B) = +1$ and there is  no $B' \in \B_{ \ell}(x)$ has both $B' \subset B$ and $y(B') = -1$.
\item A similar condition determines whether $\PL_\ell(x)$ contains $-1$.
\item
Given this set of possible labels of $x$, we assign a label
$$ y_\ell(x)
=
\left\{
\begin{array}{cl}
+1 & \mbox{if $+1 \in \PL_\ell(x)$  and $-1 \notin \PL_\ell(x)$}\\
-1 & \mbox{if $-1 \in \PL_\ell(x)$  and $+1 \notin \PL_\ell(x)$}\\
0 & \mbox{if $\PL_\ell(x) = \{0\}$} \\
! & \mbox{if $+1 \in \PL_\ell(x)$  and $-1 \in \PL_\ell(x)$}
\end{array}
\right.
$$
\item Define $U_\ell = \{x \in X: y_{\ell-1}(x) = !\}$
\item Define the set of query balls 
$$Q_l= \left\{B \in \B_l  |   B \cap U_{l-1} \neq \emptyset \right\}$$
And define the focus set to be the union of the query balls:
$$F_l = \bigcup_{B \in Q_l} B$$
\end{itemize}

\item Implementing the oracle. Let $P_\ell = \Pr[F_\ell]$
Our hypothesis is that, if the set of balls has $d$, if we we select $M_\ell = C (\log d \log \delta^{-1})(n k 2^\ell) P_\ell$ independent points in $F_\ell$ then with probability at least $1-\delta$ each ball contains at least $k$ samples.

\item Deterministic algorithm\\
Set $F_1 = X$  \\
For $l=1,\ldots, l_{max}$
\begin{itemize}
\item {\em Focused Query:} Query $M_\ell$ from $F_\ell$
\item {\em Uniform query:} Query $M_\ell$ from $X$
\item {\em Back-filling:} Consider a "hindsight known unknown" to be a set 
$U_m \neq \emptyset$ for $m<\ell$ that was not identified in iteration $\ell-1$. Then focused sampling is added for levels $m+1$ till $\ell$. (needs clearer explanation)
\end{itemize}
\end{enumerate}

\section{Sketch of analysis}

Consider the number of focused queries made at level $\ell$: $M_\ell = C (\log d \log \delta^{-1})(n k 2^\ell) P_\ell$.
Denote $N_\ell = 2C (\log d \log \delta^{-1})(n k 2^\ell)$. Then $M_{\ell} = N_\ell P_\ell$. is the number of samples needed to guarantee that the balls in $Q_\ell$ all contain at least $k$ samples. 
If we replace $P_\ell$ with $1$ we get $N_l$: the number of uniform queries needed to get reliable labels for all 
of the balls in $\B_\ell$.

Fix a point $x\in X$, define the history $H(x)$ to be as binary sequence of length $l_{max}$ where each element is either 
$C$ for {\em consistent} or $I$ for {\em Inconsistent}.  We use $H_i(x)$ to indicate the $i$'th element in $H(x)$.
$H_i(x)=I$ if and only if $x \in F_\ell$.

We now characterize the convergence of the predictions on $x$.  We say that $c(x)$ is the {\em convergence point for $x$} if $j=c(x)$ is the smallest integer such that for all $i\geq j$, $H_i(x)=C$. we say that $d(x)$ is the {\em final disagreement point} if $j=d(x)$ is the smallest integer such that for all $j \geq i > c(x)$, $H_j(x)=I$

Our analysis is based on the following intuition. For levels $1,\ldots,d(\ell)-1$ we need to use uniform sampling while for levels $d(\ell),\ldots c(\ell)-1$ we use focused sampling.

As a result the number of samples needed is 
$$ 2\left( \sum_{\ell=1}^{d(x)-1} N_{\ell} + \sum_{\ell=d(x)}^{c(x)-1} N_\ell P_\ell \right) $$ while the number of queries needed  using standard uniform queries is  
$$ \sum_{\ell=1}^{c(x)-1} N_{\ell} $$
Easy examples are ones where the $d(x)$ and $c(x)$ are small and $P_\ell$ is small for $d(x)\geq \ell \geq c(x)$. Note that the sequence $P_1,P_2$ is global and does not depend on $x$. The dependence on $x$ comes in through $d(x)$ and $c(x)$
\pagebreak
\section{1D example}
I believe that for each point $x$ there are two critical levels. In the following I say that balls are "in agreement" if there is a sign $s \in \pm 1$ such that each ball is labeled either $s$ or $?$
\begin{itemize}
    \item The convergence level: all balls at this or higher level are in agreement with the polarity same $s$.
    \item The last unknown-unknown level: the highest level in which there is agreement and the label is wrong.
\end{itemize}
Basically, we rely on uniform sampling until we pass the last unknown-unknown level. We rely on focused sampling from that level until convergence.

In the figure above I sketch a simple case where there are two thresholds and $\eta$ is either $1/2-\gamma$ or $1/2+\gamma$. In this case the convergence radius is $d/2$ and the last UU is radii above $2W$

\begin{figure}[t]
\begin{center}
\includegraphics[width=4in]{1Dbump.jpg}
\end{center}
\label{fig:cases}
\end{figure}

@@

\subsection{The algorithm}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.4in}
\begin{itemize}
\item Initialize uncertainty regions at all levels:
\begin{itemize}
\item $U_0 = X$
\item $U_\ell = \emptyset$ for $\ell \geq 1$
\end{itemize}

\vspace{.1in}
\item Repeat:
\begin{itemize}
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}
\item Let $\ell'$ be the smallest such level
\item {\tt Focused-query}($\ell', U_{\ell'}$)
\end{itemize}
\item {\tt Background-query}
\item Update $\yh_{\ell}$ values, $\ell \geq 0$, as per equation (\ref{eq:provisional-label})
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
\end{itemize}

\begin{figure}[h!]
\framebox{
\begin{minipage}[t]{6.4in}

\vspace{.1in}
\emph{Initialization:}
\begin{itemize}
\item Set $Q = \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{unif}[0,1]$
\end{itemize}

@@

\vspace{.1in}
\begin{itemize}
\item Query the next unlabeled point in $X \setminus Q$, ordered by $T_z$ values, and add to $Q$
\end{itemize}

\end{minipage}}
\caption{The sampling procedures. Each point $x \in X$ has an associated value $T_x$ chosen uniformly at random from $[0,1]$. This smaller this value, the earlier this point is likely to be queried. The focused querying procedure also uses threshold $\tau_\ell = 2^{\ell+2}k/n$.} 
\label{alg:sampling}
\end{figure}

@@

Our main deviation bound is the following, a consequence of Lemma~\ref{lemma:large-deviation-discrete} in the appendix.
\begin{thm}
Suppose $k \geq 2c_1^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c_1}{\sqrt{k}}.$$
\end{enumerate}
\label{thm:large-deviation-bounds}
\end{thm}
\begin{proof}
These follow immediately by applying Lemma~\ref{lemma:large-deviation-discrete} to each $B \in \B$ and taking a union bound. 
\end{proof}

@@

@@

\caption{This figure provides some examples for the concept of mind changes. Here we consider true probabilities, rather than estimates. The bounding rectangle defines the input space $X$. The blue region indicate $x$'s where the probability of +1 is larger than that of $-1$, the red region indicates the revers polarity. More specifically, we assume that in the blue region the probability of +1 is 0.6, while in the red region the probability of +1 is 0.4.
The black dots correspond to points on which we want to make a prediction. We consider all circles centered at each point. The {\em bias} of  circle is the probability of +1 conditioned on the point being in the circle. As labels are accumulated, the bias of smaller and smaller circles can be determined. For any point there is a large enough circle that contains the whole input space. We assume that the overall bias is negative, hence for a sufficiently large circle the bias is negative. This is indicated by the dotted circle which is always the largest circle. The solid line circles are the largest circles with the polarity indicated by the color of the line. The smallest circle corresponds to the last mind change which we also call the convergence circle.
A point is considered "easy" if its convergence circle is large, because that means that a small number of queries is required to find the correct label.
Following that definition we get that the order of points from the easiest to the hardest is {\bf a,b,c,e,d}. Comparing e and d we see that d has a smaller convergence circle, and one mind change, which (e) has a larger convergence circle, but has 4 mind changes.
}

@@

\iffalse
\begin{figure}
\begin{center}
\includegraphics[width=6in]{IMG_0840.jpg}

%\includegraphics[width=1.5in]{boundary3.pdf}
%\hskip1in
%\includegraphics[width=1.75in]{boundary4.pdf}
\end{center}
\caption{\emph{Left:} In this picture, it could be the case that every ball $B \in \B(x)$ has $\eta(B) \geq 1/2$, in which case $L_s(x) = 0$. But these balls may not have significant positive bias until a higher level $L_f(x)$ (which includes the shown ball). \emph{Right:} This is a situation in which, for small $\ell$, balls $B \in \B_\ell(x)$ might have $\eta(B) < 1/2 - \gamma$. But for $\ell \geq L_s(x)$, we get balls (such as the outer ball) with a positive bias, and for $\ell \geq L_f(x)$ (such as the inner ball), a significant positive bias.}
\label{fig:critical-interval}
\end{figure}
\fi
